\PYGEOMETRY{}: a Python library for the creation of Monte Carlo radiation transport physical geometries

### 2020-12-17 public/paper/pyg4ometry.tex
Stuart first
add laurie's last fixes including citing and mentioning Cython

\author[a]{S.D. Walker}
\author[a]{A. Abramov}
\author[a]{L.J. Nevay}
\author[a]{W. Shields}
\author[a]{S.T. Boogert\corref{author}}

@@

%% main text
\section{Introduction} \label{sec:introduction}
There are numerous different software codes to simulate the passage of particles through material, such radiation transport (RT) programs
include MCNP~\cite{Mcnp_Werner}, FLUKA~\cite{Fluka_Ferrari,Fluka_Bohlen}, Geant3~\cite{Geant3_Brun} and Geant4~\cite{Geant4_Agostinelli}.
All these codes are based on the Monte Carlo technique but each code either has a particular specialism, simulation methodology or target user community.
 Monte Carlo RT (MCRT) simulations have diverse uses including shielding calculations for radiological protection, detector performance, medical
imaging and therapy, and space radiation environment simulations. A fundamental requirement of all of the codes is to supply a
computer-readable description of the  physical three-dimensional geometry that  the particles are passing through.  The creation of geometry files is
typically a very time-consuming activity and the simulation validity and performance is directly dependent on the quality of the geometry. There is no
standard geometry format used across MCRT codes, with each code often using its own unique format. A user will typically not have geometry in both FLUKA and Geant4 for example. A geometry system
that enables the conversion between files prepared for different codes will allow for cross-checks of the physics processes in different particle transport
 codes.  The file formats used for geometry are generally focused
 on the computational efficiency of particle tracking algorithms  and not ease of use. In addition to the creation of geometry files for RT programs, usually
 computer-aided design (CAD) files exist for systems which need to be simulated. The fundamental geometric representations in CAD files are usually not
 amenable to MCRT programs.  For these reasons it is advantageous to create a software tool that allows particle transport code users to rapidly develop
 error-free geometry files, convert between common MCRT geometry formats and load CAD models.

@@

\subsection{Tessellation of solids (meshing)}
Creating a uniform three-dimensional mesh description of all solids (including Booleans) is exceptionally useful for visualisation and other algorithms, such as overlap
detection. For each Geant4 solid instance a triangular tessellated vertex-face mesh is generated and cached. This mesh is then used to determine the extent
of placed instances of geometry (physical volumes) and meshes for CSG-derived solids. CSG mesh calculations are performed using a Binary Space Partitioning
(BSP) tree technique in pure Python~\cite{pycsg} or via the Computational Geometry Algorithms Library (CGAL) surface meshes~\cite{cgal:bsmf-sm-20b} in C++. In
general, the CGAL implementation is one to two orders of magnitude faster than the pure Python CSG implementation and must be used for large geometries.
Triangular meshes based on CSG operations involving curved surfaces often contain large numbers of triangles. Before meshes are visualised or written to file
various polygon mesh algorithms  from CGAL~\cite{cgal:lty-pmp-20b} can be employed to give the meshes more desirable features.

>>>>

\author[a]{S.~D. Walker}
\author[a]{A. Abramov}
\author[a]{L.~J. Nevay}
\author[a]{W. Shields}
\author[a]{S.~T. Boogert\corref{author}}

@@

%% main text
\section{Introduction} \label{sec:introduction}
There are numerous different software codes to simulate the passage of particles through material, such radiation transport (RT) programs
include MCNP~\cite{Mcnp_Werner}, FLUKA~\cite{Fluka_Ferrari,Fluka_Bohlen}, Geant3~\cite{Geant3_Brun} and Geant4~\cite{Geant4_Agostinelli}.
All these codes are based on the Monte Carlo technique but each code either has a particular specialism, simulation methodology or target user community.
 Monte Carlo RT (MCRT) simulations have diverse uses including shielding calculations for radiological protection, detector performance, medical
imaging and therapy, and space radiation environment simulations. A fundamental requirement of all of the codes is to supply a
computer-readable description of the  physical three-dimensional geometry that  the particles are passing through.  The creation of geometry files is
typically a very time-consuming activity and the simulation validity and performance is directly dependent on the quality of the geometry. There is no
standard geometry format used across MCRT codes, with each code often using its own unique format. A user will typically not have geometry in both FLUKA and Geant4 for example. A geometry system
that enables the conversion between files prepared for different codes will allow for cross-checks of the physics processes in different particle transport
 codes.  The file formats used for geometry are generally focused
 on the computational efficiency of particle tracking algorithms  and not ease of preparation. In addition to the creation of geometry files for RT programs, usually
 computer-aided design (CAD) files exist for systems which need to be simulated. The fundamental geometric representations in CAD files are usually not
 amenable to MCRT programs.  For these reasons it is advantageous to create a software tool that allows particle transport code users to rapidly develop
 error-free geometry files, convert between common MCRT geometry formats and load CAD models.

@@

\subsection{Tessellation of solids (meshing)}
Creating a uniform three-dimensional mesh description of all solids (including Booleans) is exceptionally useful for visualisation and other algorithms, such as overlap
detection. For each Geant4 solid instance a triangular tessellated vertex-face mesh is generated and cached. This mesh is then used to determine the extent
of placed instances of geometry (physical volumes) and meshes for CSG-derived solids. CSG mesh calculations are performed using a Binary Space Partitioning
(BSP) tree technique in pure Python~\cite{pycsg} or via the Computational
Geometry Algorithms Library (CGAL) surface meshes~\cite{cgal:bsmf-sm-20b}
in C++. The pure Python CSG backend is further accelerated by compiling it
to C using Cython\cite{cython}.
For maximum flexibility either backend can be used, but in general, the CGAL implementation is one to two orders of magnitude faster than the Cythonized CSG implementation and must be used for large geometries.
Triangular meshes based on CSG operations involving curved surfaces often contain large numbers of triangles. Before meshes are visualised or written to file
various polygon mesh algorithms  from CGAL~\cite{cgal:lty-pmp-20b} can be employed to give the meshes more desirable features.
